---
title: "TURF Analysis Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TURF Analysis Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## What is TURF?

TURF (Total Unduplicated Reach and Frequency) finds the combination of items that reaches the most people. Given a set of binary survey items (e.g., "would you buy this flavor?"), TURF answers: **which N items should we offer to maximize the number of people who select at least one?**

The `turf` package provides two main functions:

- `turf_best_combo()` — runs the analysis
- `turf_write()` — writes results to an interactive Excel workbook

## Example data

The package includes synthetic ice cream survey data for testing:

```{r}
library(turf)

# 1000 respondents, 30 ice cream flavors
head(example_data_ice_cream)

# Variable-to-label mapping
example_data_ice_cream_dictionary
```

---

## turf_best_combo()

Computes TURF for every n-item combination from a binary (0/1) data frame.

### Basic usage

```{r}
results <- turf_best_combo(
  df   = example_data_ice_cream,
  vars = example_data_ice_cream_dictionary$variable,
  n    = 1:5
)
```

### Parameters

#### `df` (required)

A data frame containing the binary item columns, and optionally weight and subgroup columns. Each item column should be 0/1 where 1 means the respondent selected that item.

#### `vars` (required)

Character vector of column names in `df` that are the binary items to analyze. Only these columns are evaluated — everything else (weights, subgroups, IDs) is ignored by the engine.

```{r}
vars <- c("ic_vanilla", "ic_chocolate", "ic_strawberry", "ic_mint_chip")
```

#### `n`

Integer or integer vector specifying the combination size(s) to evaluate.

- Single value (e.g., `n = 3`): returns a single tibble of all 3-item combos
- Vector (e.g., `n = 1:5`): returns a named list of tibbles (`$n_1`, `$n_2`, ..., `$n_5`)

Default: `1:5`

```{r}
# Just 3-item combos
results_3 <- turf_best_combo(df, vars, n = 3)

# Combos of size 1 through 10
results_all <- turf_best_combo(df, vars, n = 1:10)
results_all$n_5  # access the 5-item results
```

#### `subgroups`

Character vector of column names in `df`. Each should be a binary (0/1) indicator. TURF runs separately on respondents where each column equals 1.
Use a column of all 1s for the total sample (e.g., `"Total"`).

When subgroups are provided, results are returned as a named list by subgroup, where each element contains the tibble(s) for that subgroup.

```{r}
results <- turf_best_combo(
  df        = example_data_ice_cream,
  vars      = example_data_ice_cream_dictionary$variable,
  n         = 1:5,
  subgroups = c("Total", "Gen_Z", "Millennials", "Gen_X")
)

# Access Gen_Z results for 3-item combos
results$Gen_Z$n_3
```

#### `labels`

A data frame with columns `variable` and `label` (e.g., from a data dictionary). When provided, label columns (`label_1`, `label_2`, ...) are interleaved next to each item column in the output.

```{r}
results <- turf_best_combo(
  df     = example_data_ice_cream,
  vars   = example_data_ice_cream_dictionary$variable,
  n      = 3,
  labels = example_data_ice_cream_dictionary
)
# Output columns: item_1, label_1, item_2, label_2, item_3, label_3, reach_n, ...
```

The function also accepts:

- A named character vector (names = variable names, values = labels)
- Column name variations like `var`, `vars`, `lab`, `labs` are auto-detected

#### `weight`

String. Column name in `df` containing respondent weights. When provided, weighted reach and frequency columns are appended alongside the unweighted columns:

| Column | Description |
|--------|-------------|
| `reach_n` | Unweighted count reached |
| `reach_pct` | Unweighted percent reached |
| `freq_avg` | Unweighted average frequency |
| `w_reach_n` | Weighted count reached |
| `w_reach_pct` | Weighted percent reached |
| `w_freq_avg` | Weighted average frequency |

```{r}
results <- turf_best_combo(df, vars, n = 3, weight = "weight")
```

#### `parallel`

Logical. Default `TRUE`. Enables parallel processing via `future`/`furrr` when the combo space is large enough. When subgroups are provided and the combo space is small, subgroups themselves run in parallel.

Requires the `future` and `furrr` packages (listed in Suggests).

#### `max_brute` and `max_parallel`

These thresholds control which search strategy is used:

| Combo count | Strategy |
|-------------|----------|
| <= `max_brute` | Brute-force, single core |
| `max_brute` to `max_parallel` | Parallel brute-force |
| > `max_parallel` | Hybrid (greedy + random + shortlist) |

Defaults: `max_brute = 150000`, `max_parallel = 150001`

With the defaults, brute-force is used up to 150k combos and hybrid kicks in above that.

#### `n_random`

Integer. Number of random combos to sample in the hybrid strategy. Default `100000`. Larger values improve coverage of the combo space at the cost of runtime.

#### `n_shortlist`

Integer. Number of top items (by individual reach + greedy path) to include in the shortlist brute-force pass of the hybrid strategy. Default `20`.

#### `seed`

Integer. Seed for reproducibility of random sampling in the hybrid strategy. Default `1`.

#### `estimate_time`

Logical. If `TRUE`, runs a quick benchmark and prints an estimated runtime table without actually running the analysis. Returns the estimate invisibly.

```{r}
turf_best_combo(df, vars, n = 1:10, estimate_time = TRUE)
#> Benchmark: 12,345 combos/sec (n=5, 1,000 respondents)
#> # A tibble: 10 x 5
#>        n   combos strategy est_sec est_min
#>    <int>    <dbl> <chr>      <dbl>   <dbl>
#>  1     1       30 brute        0.0     0.0
#>  2     2      435 brute        0.0     0.0
#>  ...
```

### Return value

Each tibble in the output is sorted by reach (descending) and contains:

| Column | Description |
|--------|-------------|
| `item_1` ... `item_n` | Variable name for each item in the combo |
| `label_1` ... `label_n` | (If labels provided) Human-readable label |
| `reach_n` | Number of respondents reached by this combo |
| `reach_pct` | Percent of respondents reached |
| `freq_avg` | Average number of items selected among reached respondents |
| `w_reach_n` | (If weighted) Weighted count reached |
| `w_reach_pct` | (If weighted) Weighted percent reached |
| `w_freq_avg` | (If weighted) Weighted average frequency |

### Search strategies

The function automatically selects the best strategy based on the number of combinations:

**Brute-force** — evaluates every possible combination. Exact results, but only feasible for smaller combo spaces.

**Parallel brute-force** — same as brute-force but distributed across CPU cores via `future`/`furrr`.

**Hybrid** — for very large combo spaces (e.g., 40 items choose 5 = 658k combos), uses three complementary approaches:

1. **Greedy**: stepwise selection of the best item at each step. Near-optimal and fast.
2. **Shortlist**: takes the union of greedy-path items and top individual-reach items, then brute-forces all combos within that reduced set. Exhaustively covers the most promising region.
3. **Random**: samples broadly across the full space as a hedge against greedy/shortlist missing an unexpected combo.

---

## turf_write()

Writes TURF results to an interactive macro-enabled Excel workbook (.xlsm) with:

- **Dashboard** sheet: dropdown controls, greedy build chart, item toggle panel
- **Best Combos** sheet: combo results table with sorting/filtering controls
- Hidden data sheets for VBA to read from

### Basic usage

```{r}
turf_write(
  best_combo_results = results,
  raw                = example_data_ice_cream,
  vars               = example_data_ice_cream_dictionary$variable,
  subgroups          = c("Total", "Gen_Z", "Millennials", "Gen_X"),
  weight             = "weight",
  labels             = example_data_ice_cream_dictionary,
  file_name          = "my_turf_analysis",
  project_name       = "Ice Cream Study (#1234567)"
)
```

### Parameters

#### `best_combo_results`

Output from `turf_best_combo()`. Accepts any of the 4 return shapes:

- Single tibble (one n, no subgroups)
- Named list by n (`$n_1`, `$n_2`, ...)
- Named list by subgroup (each containing a tibble)
- Nested list (subgroup > n)

If `NULL`, the Best Combos sheet is created with dummy controls and hidden. The Dashboard and greedy functionality still work without it.

#### `raw` (required)

The original respondent-level data frame. Must contain the item columns, weight column (if used), and subgroup columns (if used). This is written to a hidden `_raw` sheet so the VBA macros can compute greedy results interactively.

#### `vars` (required)

Character vector. Same item variable names passed to `turf_best_combo()`.

#### `subgroups`

Character vector. Subgroup column names in `raw` (binary 0/1). If `NULL`, a "Total" subgroup is created automatically.

#### `weight`

String. Weight column name in `raw`. If `NULL`, all respondents are equally weighted and the weighted controls are hidden in the workbook.

#### `respondent_id`

String. Column name for respondent IDs. If `NULL`, row numbers are used. Written to the hidden `_raw` sheet.

#### `labels`

Same as `turf_best_combo()` — a data frame with `variable`/`label` columns, or a named character vector.

#### `top`

Integer. Maximum number of combos to write per subgroup x combo size. Default `50000`. Larger values increase file size.

#### `file_name`

String. Output file name without extension. Default `"TURF_Analysis"`. The extension (`.xlsm` or `.xlsx`) is determined automatically based on whether the VBA template is found.

#### `project_name`

String. Displayed in row 3 of the Dashboard and Best Combos sheets. Default `"Project Name - (#xxxxxxx)"`.

#### `where`

String. Output directory. Default `getwd()`.

#### `template`

String. Path to the VBA template `.xlsm` file. By default, the function looks for the bundled template in the `turf` package (`inst/turf_template.xlsm`), then falls back to `turf_template.xlsm` in the working directory.

If the template is found, the output is `.xlsm` with full VBA interactivity. Without it, the output is a static `.xlsx`.

#### `sig_threshold`

Numeric. P-value threshold for green significance coloring in the greedy results. Default `0.10`.

#### `marginal_threshold`

Numeric. P-value threshold for orange marginal significance. Default `0.20`. Values above this threshold are colored red.

#### `very_hidden`

Logical. If `TRUE`, hidden sheets are set to Excel's "veryHidden" state (only accessible via VBA editor). Default `FALSE`.

#### `return_location`

Logical. If `TRUE` (default), returns the file path invisibly.

### Workbook structure

The generated workbook contains:

| Sheet | Visible | Purpose |
|-------|---------|---------|
| Dashboard | Yes | Greedy TURF chart, results table, item controls |
| Best Combos | Yes | Combo results with sorting/filtering |
| Greedy | Hidden | Greedy results + chart staging data |
| `_raw` | Hidden | Respondent-level data for VBA computation |
| `_config` | Hidden | Settings, variable/label mapping, base sizes |
| `_controls` | Hidden | Dropdown state (single source of truth) |
| `_best_combo_charts` | Hidden | Chart staging data |
| `d_s1_2`, `d_s1_3`, ... | Hidden | Combo data per subgroup x combo size |

The Dashboard sheet includes interactive dropdowns for subgroup, optimize metric (Reach/Freq), chart label format, and weighted toggle. The VBA macros read from the hidden sheets and update the visible sheets when dropdown values change.

---

## Full example

```{r}
library(turf)

# Step 1: Run TURF
turf_results <- turf_best_combo(
  df        = example_data_ice_cream,
  vars      = example_data_ice_cream_dictionary$variable,
  n         = 1:5,
  subgroups = c("Total", "Gen_Z", "Millennials", "Gen_X"),
  labels    = example_data_ice_cream_dictionary,
  weight    = "weight"
)

# Step 2: Write to Excel
turf_write(
  best_combo_results = turf_results,
  raw                = example_data_ice_cream,
  vars               = example_data_ice_cream_dictionary$variable,
  subgroups          = c("Total", "Gen_Z", "Millennials", "Gen_X"),
  weight             = "weight",
  labels             = example_data_ice_cream_dictionary,
  top                = 1000,
  file_name          = "example_ice_cream_turf",
  project_name       = "Ice Cream Study (#1234567)"
)
```
